The impact of cosmic variance on simulating weak lensing surveys

### 2014-08-26 Writeup/paper_v2.tex
References from Mandelbaum et. al 2012 added

\begin{keywords}
 Gravitational lensing: weak, Cosmology: Large-scale structure of Universe, Galaxies: evolution {\bf ?} \rachel{Are these from the MNRAS list?}
\end{keywords}

@@

\rachel{DES is far from the only survey; since I'm in HSC and the work was supported by Euclid project, it seems strange to only mention DES.  I recommend taking one of my papers (e.g., the one on image simulations from 2012) and lifting the entire list of references to surveys directly from there.}
Major surveys such as the Dark Energy Survey (DES) are planned to be carried out over the next decade or two to make measurements that constrain the nature of dark energy.
Weak gravitational lensing probes the nature of dark energy through both the growth of structure in the universe with redshift and the dependence of distances on the expansion rate or the Hubble parameter.
However, most of the matter in the universe is made up of dark matter which cannot be detected by most of the standard astronomical/cosmological probes. Weak lensing happens to be the cleanest probe for this purpose.
As light from the background sources passes through the foreground cluster of galaxies and dark matter profiles, the shapes get distorted. Weak lensing also causes magnification in addition to introducing a shear.
%Weak lensing effects need to be removed from the Cosmic Microwave Background (CMB) analysis to understand the initial conditions of the universe.

@@

%The structure of the paper is as follows. In \S\ref{S:data}, we briefly describe the COSMOS catalog that we use for our analysis.
In this paper, we show that these large scale structure effects mentioned are indeed present in the data. The rest of the paper is structured as follows.
In \S\ref{sub:overdensities}, we explain how we identify overdensities and underdensities, volume-limit our data in \S\ref{sub:volumelimiting} and in \S\ref{sub:axisratio},
 we explain how we obtain axis-ratios from the images of the galaxies. Finally, we present our results in \S\ref{S:results} and end with conclusions in \S\ref{S:conclusion}.
\section{Data}
\label{S:data}
%In this analysis, we consider the COSMOS sample of galaxies. 
COSMOS~\citep{COSMOS_overview, COSMOS_generic, COSMOS_Alexie} is a flux-limited, narrow deep field survey covering a contiguous area of $1.64 \text{ deg}^2$ of sky, with images taken using the Advanced Camera for Surveys (ACS) Wide Field Channel (WFC)
in the Hubble Space Telescope (HST). Precise shape measurements, when compared to ground-based surveys, can be made since the full width half-maximum (FWHM) of the point-spread function (PSF)
is $0.12''$. High resolution images taken through the wide F814W filter (broad \emph{I}), after correcting for the PSF, have led to collection of postage stamp images\footnote{http://irsa.ipac.caltech.edu/data/COSMOS/images/galaxy-postage-stamps/} for the GREAT3 challenge~\citep{great3}, using the procedure described in \cite{SHERA}.
Parametric models to most of these galaxies including \sersicn profile fits, 
2 component bulge+disk fits, axis ratios etc were done for GREAT3 challenge~\citep{great3} using the method described in \cite{Claire_Fits}. This is described briefly in Sec.~\ref{sub:axisratio}.In addition to the ACS/WFC (F814W) imaging, the COSMOS field has also been imaged by Subaru Suprime-Cam ($B_j, V_j, g^+,r^+,i^+,z^+,$ NB816), the
Canada-French Hawaii Telescope (CFHT;$u^*,i^*$) and the KPNO/CTIO ($K_s$) .

@@

 \item \texttt{GOOD\_ZPHOT\_SOURCE=1}: This cut requires that there be a good photometric redshift, which typically is equivalent to requiring that the galaxy not be located within
                                       the masked regions of the Subaru $BVIz$ imaging used for photometric redshifts, and that it have a successful match against an object in the Subaru imaging
                                       {\bf copied from Mandelbaum 2011.}
\end{enumerate}
\rachel{The cuts you just gave are really basic or fundamental ones.  Those should go before any discussion of more complicated derived quantities like redshifts, stellar mass, Sersic fits, etc.}

@@

\subsection{Volume limiting}
\label{sub:volumelimiting}
COSMOS is a flux-limited survey and is therefore affected by Malmquist bias, i.e. at higher redshifts, the brighter galaxies are preferentially observed.
Our analysis involves comparing galaxies in different redshift slices to find significant differences in morphology, if any, so with a flux-limited sample, we would be comparing only the bright galaxies at high redshifts with
bright and faint galaxies at low redshifts. For a fair comparison, we must restrict ourselves to bright galaxies at all redshifts and this is acheived by volume-limiting the sample.
%\rachel{Explain why you are doing this.  Before saying {\em how} you
%  generate a volume-limited sample, explain {\em why} you are
%  generating one.}
We generate, using the method below, a volume-limited sample that is complete upto $z=1$ by applying a cut on luminosity such that only galaxies instrinsically brighter than a certain threshold is considered. This threshold is set on $K$-corrected $i$-band magnitudes, denoted as $M_I$.
\rachel{Is it $i$ band, or $I$ band?  Check which one and then be
  self-consistent throughout the text.}
Since the parent sample contains fainter galaxies, upto F814W=25.2, we compare the distribution of the F814W=23.5 sample with the samples containing fainter galaxies for high redshift bins, to see where the sample is no longer complete.
\begin{figure}
 \centering
 a) \includegraphics[width=\columnwidth]{volume_limiting_pdf}
 b) \includegraphics[width=\columnwidth]{volume_limiting_cdf}
 \label{fig:volume_limiting_dist}
 \caption{ ... }
\end{figure}

@@

%\begin{thebibliography}
\bibliography{ref}
%\end{thebibliography}

@@

\end{document}

>>>>

\begin{keywords}
 Gravitational lensing: weak, Cosmology: Large-scale structure of Universe, Galaxies: evolution {\bf ?} \rachel{Are these from the MNRAS list?} \arun{Yes.}
\end{keywords}

@@

\rachel{DES is far from the only survey; since I'm in HSC and the work was supported by Euclid project, it seems strange to only mention DES.  I recommend taking one of my papers (e.g., the one on image simulations from 2012) and lifting the entire list of references to surveys directly from there.}
Major surveys such as the Hyper Suprime-Cam \arun{How to add shortforms?}~\citep{HSC_Overview}, Dark Energy Survey~\citep{DESC}, the KIlo-Degree Survey~\citep{KIDS}, the Panoramic Survey Telescope and Rapid Response System~\citep{PanSTARRS_2010},
the Large Synoptic Survey Telescope~\citep{LSST_Book},Euclid\footnote{http://sci.esa.int/science-e/www/area/index.cfm?fareaid=102}, Wide-Field Infrared Survey Telescope~\citep{WFIRST_Final}
are planned to be carried out over the next decade or two that will gather huge amounts of weak lensing data that can be used to constrain cosmological parameters more precisely than ever.
Weak gravitational lensing probes the nature of dark energy through both the growth of structure in the universe with redshift and the dependence of distances on the expansion rate or the Hubble parameter.
However, most of the matter in the universe is made up of dark matter which cannot be detected by most of the standard astronomical/cosmological probes. Weak lensing happens to be the cleanest probe for this purpose.
As light from the background sources passes through the foreground cluster of galaxies and dark matter profiles, the shapes get distorted. Weak lensing also causes magnification in addition to introducing a shear.
%Weak lensing effects need to be removed from the Cosmic Microwave Background (CMB) analysis to understand the initial conditions of the universe.

@@

%The structure of the paper is as follows. In \S\ref{S:data}, we briefly describe the COSMOS catalog that we use for our analysis.
In this paper, we show that these large scale structure effects mentioned are indeed present in the data. The rest of the paper is structured as follows.
In \S\ref{sub:overdensities}, we explain how we identify overdensities and underdensities, volume-limit our data in \S\ref{sub:volumelimiting} and in \S\ref{sub:axisratio},
 we explain how we obtain axis-ratios from the images of the galaxies. Finally, we present our results in \S\ref{S:results} and end with conclusions in \S\ref{S:conclusion}.
\section{Data}
\label{S:data}
%In this analysis, we consider the COSMOS sample of galaxies. 
COSMOS~\citep{COSMOS_overview, COSMOS_generic, COSMOS_Alexie} is a flux-limited, narrow deep field survey covering a contiguous area of $1.64 \text{ deg}^2$ of sky, with images taken using the Advanced Camera for Surveys (ACS) Wide Field Channel (WFC)
in the Hubble Space Telescope (HST).

 \item \texttt{GOOD\_ZPHOT\_SOURCE=1}: This cut requires that there be a good photometric redshift, which typically is equivalent to requiring that the galaxy not be located within
                                       the masked regions of the Subaru $BVIz$ imaging used for photometric redshifts, and that it have a successful match against an object in the Subaru imaging
                                       {\bf copied from Mandelbaum 2011.}
\end{enumerate}
%\rachel{The cuts you just gave are really basic or fundamental ones.  Those should go before any discussion of more complicated derived quantities like redshifts, stellar mass, Sersic fits, etc.}

Precise shape measurements, when compared to ground-based surveys, can be made since the full width half-maximum (FWHM) of the point-spread function (PSF)
is $0.12''$. High resolution images taken through the wide F814W filter (broad \emph{I}), after correcting for the PSF, have led to collection of postage stamp images\footnote{http://irsa.ipac.caltech.edu/data/COSMOS/images/galaxy-postage-stamps/} for the GREAT3 challenge~\citep{great3}, using the procedure described in \cite{SHERA}.
Parametric models to most of these galaxies including \sersicn profile fits, 
2 component bulge+disk fits, axis ratios etc were done for GREAT3 challenge~\citep{great3} using the method described in \cite{Claire_Fits}. This is described briefly in Sec.~\ref{sub:axisratio}.In addition to the ACS/WFC (F814W) imaging, the COSMOS field has also been imaged by Subaru Suprime-Cam ($B_j, V_j, g^+,r^+,i^+,z^+,$ NB816), the
Canada-French Hawaii Telescope (CFHT;$u^*,i^*$) and the KPNO/CTIO ($K_s$) .

@@

@@

\subsection{Volume limiting}
\label{sub:volumelimiting}
COSMOS is a flux-limited survey and is therefore affected by Malmquist bias, i.e. at higher redshifts, the brighter galaxies are preferentially observed.
Our analysis involves comparing galaxies in different redshift slices to find significant differences in morphology, if any, so with a flux-limited sample, we would be comparing only the bright galaxies at high redshifts with
bright and faint galaxies at low redshifts. For a fair comparison, we must restrict ourselves to bright galaxies at all redshifts and this is acheived by volume-limiting the sample.
%\rachel{Explain why you are doing this.  Before saying {\em how} you
%  generate a volume-limited sample, explain {\em why} you are
%  generating one.}
We generate, using the method below, a volume-limited sample that is complete upto $z=1$ by applying a cut on luminosity such that only galaxies instrinsically brighter than a certain threshold is considered. This threshold is set on $K$-corrected $i$-band magnitudes, denoted as $M_I$.
\rachel{Is it $i$ band, or $I$ band?  Check which one and then be
  self-consistent throughout the text.}
Since the parent sample contains fainter galaxies, upto F814W=25.2, we compare the distribution of the F814W=23.5 sample with the samples containing fainter galaxies for high redshift bins, to see where the sample is no longer complete.
% \begin{figure}
%  \centering
%  a) \includegraphics[width=\columnwidth]{volume_limiting_pdf}
%  b) \includegraphics[width=\columnwidth]{volume_limiting_cdf}
%  \label{fig:volume_limiting_dist}
%  \caption{ ... }
% \end{figure}

@@

@@

%\begin{thebibliography}
\bibliography{ref}
%\end{thebibliography}
\end{document}
